MCC Novelty Kids

### 2019-01-29 papers/cogsci2019/cogsci2019.tex
add saves to cogsci folder

\begin{abstract}
Pragmatic inferences are an integral part of language learning and
comprehension. To recover the intended meaning of an utterance,
listeners need to balance and integrate different sources of contextual
information. In a series of experiments, we studied how listeners
integrate general expectations about speakers with expectations specific
to their interactional history with a particular speaker. We used a
Bayesian pragmatics model to formalize the integration process. In
Experiments 1 and 2, we replicated previous findings showing that
listeners make inferences based on speaker-general and speaker-specific
expectations. We then used the empirical measurements from these
experiments to generate model predictions about how the two kinds of
expectations should be integrated, which we tested in Experiment 3.
Experiment 4 replicated and extended Experiment 3 to a broader set of
conditions. In both experiments, listeners based their inferences on
both types of expectations. We found that model performance was also
consistent with this finding; with better fit for a model which
incorporated both general- and specific-information compared to
baselines incorporating only one type. Listeners flexibly integrate
different forms of social expectations across a range of contexts, a
process which can be described using Bayesian models of pragmatic
reasoning. Our approach highlights how experiments and computational
models can be used in conjunction to inform theories about language use
and comprehension.

@@

Pragmatic inferences rest on a set of expectations that interlocutors
bring to the table when entering a communicative interaction. On the one
hand, speakers and listeners have the general expectation that their
partner communicates in an informative and relevant way (Sperber \&
Wilson, 2001). Grice (1991) summarised this expectation via the
``Cooperative Principle'': ``Make your contribution such as is required,
at the stage at which it occurs, by the accepted purpose or direction of
the talk exchange in which you are engaged.'' Importantly, the second
half of the Cooperative Principle describes a second type of
expectation: interlocutors expect each other to produce and interpret
utterances \emph{in light of the shared common ground between them} (H.
H. Clark, 1996). Common ground refers to bits of information that are
assumed to be shared, either because they were mentioned over the course
of the conversation or grounded through some form of joint experience
(Bohn \& Koymen, 2018). Note that by its very nature, common ground may
vary with the individuals involved in a particular conversation.

@@

Experiments were conducted online using Amazon's Mechanical Turk. Figure
\ref{fig:design} provides a schematic overview of the setup and
experimental procedures. The instructions informed participants that
they would see a number of animal characters asking for novel toys.
Participants were asked to identify the toy being requested by a
particular animal. The basic layout involved two tables with toys on
them, located left and right of a little hill, on which the animal was
standing. For each animal, we recorded a set of utterances (one speaker
per animal) that were used throughout the experiments to provide
information and make requests. At test, toys were requested using the
following utterance: ``Oh cool, there is a {[}non-word{]} on the table,
how neat, can you give me the {[}non-word{]}?''. Participants responded
by clicking on one of the toys. Each experiment started with two
training trials in which animals requested familiar objects (car and
ball).

@@

\caption[Schematic experimental procedure]{Schematic experimental procedure. In all conditions, at test (bottom), the speaker ambiguously requested an object using a non-word (e.g. “dax”). Participants clicked on the object they thought the speaker referred to. Informativeness (Experiment 1, left) translated to making one object less frequent in context. Common ground (Experiment 2, middle) was manipulated by making one object prefered by or new to the speaker. Experiment 3 (right) combined manipulations. When expressing e.g. preference for an object on a table with two objects (panel 3 from top), the respective object was temporarily enlarged. Condition shown here: preference - same speaker - incongruent. Green plus signs represent an expression of preference and red minus of dispreference (see main text for details).}\label{fig:design}
\end{figure*}
\end{CodeChunk}

@@

All participants were recruited from Amazon Mechanical Turk and had US
IP addresses. Sample size in each experiment was planned to be 120 data
points per cell. Experiment 1 had 40 participants. In the test
condition, one table contained one object of type A and the other table
contained one object of type A and one of type B (see Figure
\ref{fig:design}, left). On each trial, the animal introduced themselves
(e.g. ``Hi, I'm Dog''), turned towards the table with the two objects
and made a request. If listeners expect speakers to produce informative
utterances, they should select object B. This choice follows from the
counterfactual inference that if the (informative) speaker would have
wanted to request A, they would have turned to the table that only
contained A. On the other hand, since B is only located on the table
together with A, there was no alternative way to request B in a less
ambiguous way. In the control condition, both tables contained two
objects, one of which was randomly determined as the correct one. No
inference was therefore licenced. Each participant received three trials
in each condition for a total of six trials, presented in a randomized
order.

@@

Participants selected the less frequent object above chance in the test
condition (t(39) = 5.51, \emph{p} \textless{} .001, see Figure
\ref{fig:plotexp12}) and did so more often compared to the control
condition (genreralized linear mixed model
(GLMM\footnote{All models included random effects for subject and speaker (animal).}):\emph{\(\beta\)}
= 1.49, se = 0.5 \emph{p} = .003). This result replicates earlier work
(Frank \& Goodman, 2014) and is consistent with the hypothesis that
listeners expect speakers to communicate in an informative way.

@@

\caption[Results from Experiment 1 and 2]{Results from Experiment 1 and 2. For preference and novelty, control refers to a different speaker (see Fig. 1). Transparent dots show data from individual participants, solid dots represent condition means, error bars are 95 \% CIs. Dashed line indicates performance expected by chance.}\label{fig:plotexp12}
\end{figure}
\end{CodeChunk}

@@

We collected data from 80 participants, with 40 in each condition. In
the preference condition, each table had a different object. In the
beginning, the animal appeared on the hill and introduced themselves.
Next, they turned to one of the tables and expressed either that they
liked (``Oh wow, I really like that one'') or disliked (``Oh bleh, I
really don't like that one'') the object before turning to the other
side and expressing the respective other attitude. Then, the animal
disappeared. After a short period of time, either the same or a
different animal appeared and requested an object while facing straight
ahead (see Figure \ref{fig:design}, middle). If participants took into
account the information they gained about the speaker, they should
select the previously preferred object if the returning animal was the
same. If a different animal returned, they should choose randomly
between objects.

In the novelty condition, one table was initially empty while there was
an object on the other table (see Figure \ref{fig:design}). The animal
turned to one of the sides and commented either on the presence (``Aha,
look at that'') or the absence of an object (``Hm, nothing there'').
Next, the animal disappeared. The same animal re-appeared and the
sequence above was repeated. When the animal disappeared for the second
time, a second object appeared on the empty table while the animal was
away. Like in the preference condition, we now manipulated if the same
animal or a different one returned. In case of the same animal
returning, listeners could infer the referent of the subsequent request
by considering that one object was new to the speaker and therefore more
likely to be of interest to them. However, no such inference was
licensed when a different animal returned because both objects were
novel.

@@

Participants selected the preferred object above chance when the same
animal returned (t(39) = 29.14, \emph{p} \textless{} .001, see Figure
\ref{fig:plotexp12}) and did so more often compared to trials in which a
different animal returned (GLMM: \(\beta\) = 2.97, se = 0.7, \emph{p}
\textless{} .001). Thus, listeners inferred the referent of the
utterance by considering previous interactions with the speaker.
Interestingly, participants transferred preference to some extent from
one animal to the other and selected the preferred object above chance
when a different animal returned (t(39) = 2.7, \emph{p} = .01). In sum,
this study shows that adults make comparable inferences to children
(Saylor et al., 2009).

The novel object was selected above chance when the same animal returned
(t(39) = 6.77, \emph{p} \textless{} .001) but not when a different one
appeared (t(39) = 1.49, \emph{p} = .144, see Figure
\ref{fig:plotexp12}). Furthermore, the two conditions differed in the
expected direction (\(\beta\) = 1.7, se = 0.39, \emph{p} \textless{}
.001). Thus, like children (Akhtar et al., 1996), adults used their
prior information about the speaker to resolve ambiguity in the
utterance.

@@

\caption[Results from Experiment 3]{Results from Experiment 3. Dashed line indicates performance expected by chance. Plotting conventions are the same as in Figure 2. All conditions in which CIs do not overlap with chance line are also statistically different from chance based on two-tailed Wilcoxon tests.}\label{fig:plotexp3}
\end{figure*}
\end{CodeChunk}

In Experiment 3, we combined the expectations studied in Experiment 1
and 2 to study how listeners integrate them.

@@

A total of 121 individuals participated in the experiment. The test
situation was the same as in the test condition in Experiment 1: One
table with object of type A and the other with an object of type A and
B. Again, the animal always turned to the table with two objects and
ambiguously requested an object. In Experiment 1, the listener had no
prior information about each object. In Experiment 3, however, we
manipulated common ground expectations in the same way as in Experiment
2. For example, the animal would turn to the table with one object and
express that they don't like object A, then they would turn to the other
table and express that they like object B. Next, after quickly
disappearing, the animal would reappear, turn to the table with two
objects and make a request (see Figure \ref{fig:design}, right).

@@

To derive predictions, we used a probabilistic RSA model that simulates
a pragmatic listener reasoning about a cooperative speaker who is trying
to refer to an object (Frank \& Goodman, 2012). The speaker chooses how
to refer to the object by reasoning about a naive listener who does not
know the labels for the object (Frank \& Goodman, 2014). The conditional
probability that the listener chooses a referent given an utterance is
defined as follows: \[P_L(r_s|u)\propto P_S(u|r_s)P_S(r_s)\] Here,
\(P_S(u|r_s)\) is the likelihood that the speaker will use an utterance
u to refer to a specific referent r. It is defined in terms of a utility
function \(U_S(u;s)\) consisting of the surprisal of \(u\) for a naive
listener \(L_0\), who interprets \(u\) according its literal semantics:
\[P_S(u|r_s)\propto exp(\alpha U_S(u;s))\] The numerical strength of the
expression above depends on \(\alpha\) scalar value, α, which can be
interpreted as an indicator of how rational the speaker is in choosing
utterances (i.e.~as alpha increases, the speaker is more likely to
choose the most informative utterance).

@@

We used the results from Experiment 1 and 2 to specify \(\alpha\) as
well as \(P_S(r_s)\) in our model. We set \(\alpha\) so that a model
with uniform priors would predict the average proportion measured in
Experiment 1. The prior probability for each object was set to be the
proportion with which this object was chosen in Experiment
2\footnote{Proportions were measured when participants chose between two objects. However, in Experiment 3, three objects were involved. For each object we used the proportion measured in Experiment 2 as the prior probability.This approached spread out the absolute probability mass for each object but conserved the relative relation between objects.}.
Based on these parameter settings, we predicted the proportion with
which listeners will choose the more informative object in each of eight
unique conditions mentioned above (see also Figure \ref{fig:plotexp3}).
We compared the fit of this pragmatic model to two alternative models
using Bayes Factors (BF). The first alternative model ignored the
speaker specific expectations (uniform prior model) while the second
ignored the informativeness inference (prior only model).

@@

Results are discussed in the form of the proportion with which listeners
chose the more informative object (i.e., the object that would be the
more informative referent when only considering speaker general
information). For a comparison to chance within each unique condition
see \ref{fig:plotexp3}. Combinations of alignment and speaker identity
differed in how they influenced participants' responses (GLMM model
term: \texttt{alignment*speaker}; \(\beta\) = -2.92, se = 0.57, \emph{p}
\textless{} .001). Figure \ref{fig:plotmodelcomp}A and B shows the mean
response in each unique condition compared to the pragmatics model.
Model predictions and data were highly correlated (r = 0.96, \emph{p}
\textless{} .001).Model fit was much better in the model taking into
account both types of expectations compared to the uniform prior (BF =
2e+79) or prior only model (BF = 1.8e+34).

Interestingly, as in Experiment 2, there was a transfer of preference in
the case of speaker change. Participants were at chance in the
preference - different speaker - incongruent condition (see Figure
\ref{fig:plotexp3}). If preference would have been specific to a
particular individual, participants should have selected the less
frequent object above chance (as they did in the corresponding condition
with the novelty manipulation). Because it takes into account the
measurement from the earlier experiment, our model predicts these
results; future work might explicitly model generalization across
speakers.

@@

The strong manipulations were identical to Experiment 3 and the results
are therefore a direct replication (see Figure
\ref{fig:plotmodelcomp}C). For novelty, in the medium condition, the
animal turned to each table only once before the test. In the weak
condition, the animal only turned to the table with an object before the
test (instead of turning to and commenting on both). In the medium
condition for preference, the animal only expressed liking and did so in
a more subtle way (saying only: ``Oh, wow'' while pointing to the
object). Participants were assigned to one level of common ground
expectation and completed two test trials in each of the four conditions
(alignment x speaker change).

@@

\caption[(A) Model predictions compared to data for Experiment 3 and (B) Experiment 4]{(A) Model predictions compared to data for Experiment 3 and (B) Experiment 4. (C) Data for strong prior manipulation in Experiment 3 and 4, providing a noise ceiling for the reliability of measurements. Error bars = 95 \%  HDIs.}\label{fig:plotmodelcomp}
\end{figure*}
\end{CodeChunk}

As noted above, the strong prior condition was a direct replication of
Experiment 3. Results from the two rounds of data collection were highly
correlated (r = 0.97, \emph{p} \textless{} .001, see Figure
\ref{fig:plotmodelcomp}C). Across levels of prior manipulation, the data
from Experiment 4 were highly correlated to the corresponding model
predictions (r = 0.91, \emph{p} \textless{} .001, see Figure
\ref{fig:plotmodelcomp}B). Again, the pragmatics model provided a much
better fit compared to the flat prior (BF = 4.4e+74 ) or prior only
model (BF = 1.8e+84).

@@

Language use and learning requires balancing different types of
expectations about one's interlocutor - both expectations about how
speakers behave in general and specific expectations about how a
particular speaker might behave in a particular context. Here we used a
computational pragmatics model to predict this integration process.
Experiment 1 and 2 replicated previous studies showing that adult
listeners expect speakers to produce utterances informatively and also
with respect to common ground. We then combined the procedures from the
first two experiments to study how listeners would integrate these two
types of information. We used the results from Experiment 1 and 2 to
specify model parameters that represented the two types of expectations,
generating predictions about new behavior. Experiments 3 and 4 showed
that both types of expectations influenced listeners inferences. In
general, listener behavior was accurately described by our model,
suggesting that listeners trade-off flexibly between speaker specific
and general pragmatic expectations.

Notably, Experiment 3 also included situations in which the two
expectations were in conflict. For example, in some trials the speaker
expressed preference for object A, which was also the more frequent
(less informative) object. In these situations, a majority of
participants chose the preferred object as the referent (see also Figure
\ref{fig:plotexp3} preference--same speaker--incongruent). A simple
explanation for this pattern might be that common ground manipulations
were simply ``stronger'', corroborated by the fact they produced higher
rates of expected choice than the informativeness expectation when the
two were presented in isolation (see Figure \ref{fig:plotexp12}). In
Experiment 4, however, the medium manipulation for novelty yielded
numerically weaker results compared to the informativeness expectation
in Experiment 1, and yet participants still selected the novel object
above chance when the expectations were in conflict. Why is this?

Because common ground is represented in our model as the listener's
prior distribution, speakers can reason about it in choosing their
utterance.That is, in the mind of the listener, the speaker computes the
effect of each utterance on a naive listener with shared common ground.
Therefore, when prior interactions implicate one object as the more
likely referent, the speaker reasons that this object will be the
inferred referent of any semantically plausible utterance, even when the
same utterance would point to a different object in the absence of prior
information.

>>>>

\begin{abstract}
Pragmatic inferences are an integral part of language learning and
comprehension. To recover the intended meaning of an utterance,
listeners need to balance and integrate different sources of contextual
information. In a series of experiments, we studied how listeners
integrate general expectations about speakers with expectations specific
to their interactional history with a particular speaker. We used a
Bayesian pragmatics model to formalize the integration process. In
Experiments 1 and 2, we replicated previous findings showing that
listeners make inferences based on speaker-general and speaker-specific
expectations. We then used the empirical measurements from these
experiments to generate model predictions about how the two kinds of
expectations should be integrated, which we tested in Experiment 3.
Experiment 4 replicated and extended Experiment 3 to a broader set of
conditions. In both experiments, listeners based their inferences on
both types of expectations. We found that model performance was also
consistent with this finding; with better fit for a model which
incorporated both general and specific information compared to baselines
incorporating only one type. Listeners flexibly integrate different
forms of social expectations across a range of contexts, a process which
can be described using Bayesian models of pragmatic reasoning.

@@

Pragmatic inferences rest on a set of expectations that interlocutors
bring to the table when entering a communicative interaction. On the one
hand, speakers and listeners have the general expectation that their
partner communicates in an informative and relevant way (Sperber \&
Wilson, 2001). Grice (1991) summarised this expectation via the
\emph{Cooperative Principle}: ``Make your contribution such as is
required, at the stage at which it occurs, by the accepted purpose or
direction of the talk exchange in which you are engaged.'' Importantly,
the second half of the Cooperative Principle describes a second type of
expectation: interlocutors expect each other to produce and interpret
utterances \emph{in light of the shared common ground between them} (H.
H. Clark, 1996). Common ground refers to bits of information that are
assumed to be shared, either because they were mentioned over the course
of the conversation or grounded through some form of joint experience
(Bohn \& Koymen, 2018). Note that by its very nature, common ground may
vary with the individuals involved in a particular conversation.

@@

Experiments were conducted online using Amazon's Mechanical Turk. Fig.
\ref{fig:design} provides a schematic overview of the setup and
experimental procedures. The instructions informed participants that
they would see a number of animal characters asking for novel toys.
Participants were asked to identify the toy being requested by a
particular animal. The basic layout involved two tables with toys on
them, located left and right of a little hill, on which the animal was
standing. For each animal, we recorded a set of utterances (one speaker
per animal) that were used throughout the experiments to provide
information and make requests. At test, toys were requested using the
following utterance: ``Oh cool, there is a {[}non-word{]} on the table,
how neat, can you give me the {[}non-word{]}?''. Participants responded
by clicking on one of the toys. Each experiment started with two
training trials in which animals requested familiar objects (car and
ball).

@@

\caption[Schematic experimental procedure]{Schematic experimental procedure. In all conditions, at test (bottom), the speaker ambiguously requested an object using a non-word (e.g. “dax”). Participants clicked on the object they thought the speaker referred to. Informativeness (Experiment 1, left) translated to making one object less frequent in context. Common ground (Experiment 2, middle) was manipulated by making one object prefered by or new to the speaker. Green plus signs represent an expression of preference and red minus of dispreference (see main text for details). When expressing e.g. preference for an object on a table with two objects (panel 3 from top), the respective object was temporarily enlarged. Experiment 3 (right) combined manipulations. Condition shown here: preference - same speaker - incongruent.}\label{fig:design}
\end{figure*}
\end{CodeChunk}

@@

All participants were recruited from Amazon Mechanical Turk and had US
IP addresses. Sample size in each experiment was planned to be 120 data
points per cell. Experiment 1 had 40 participants. In the test
condition, one table contained one object of type A and the other table
contained one object of type A and one of type B (see Fig.
\ref{fig:design}, left). On each trial, the animal introduced themselves
(e.g. ``Hi, I'm Dog''), turned towards the table with the two objects
and made a request. If listeners expect speakers to produce informative
utterances, they should select object B. This choice follows from the
counterfactual inference that if the (informative) speaker would have
wanted to request A, they would have turned to the table that only
contained A. On the other hand, since B is only located on the table
together with A, there was no alternative way to request B in a less
ambiguous way. In the control condition, both tables contained two
objects, one of which was randomly determined as the correct one. No
inference was therefore licensed. Each participant received three trials
in each condition for a total of six trials, presented in a randomized
order.

@@

Participants selected the less frequent object above chance in the test
condition (t(39) = 5.51, \emph{p} \textless{} .001, see Fig.
\ref{fig:plotexp12}) and did so more often compared to the control
condition (generalized linear mixed model
(GLMM\footnote{All models included random effects for subject and speaker.}):\emph{\(\beta\)}
= 1.49, se = 0.5 \emph{p} = .003). This result replicates earlier work
(Frank \& Goodman, 2014) and is consistent with the hypothesis that
listeners expect speakers to communicate in an informative way.

@@

\caption[Results from Experiment 1 and 2]{Results from Experiment 1 and 2. For preference and novelty, control refers to a different speaker (see Fig. 1). Transparent dots show data from individual participants, solid dots represent condition means, error bars are 95\% CIs. Dashed line indicates performance expected by chance.}\label{fig:plotexp12}
\end{figure}
\end{CodeChunk}

@@

We collected data from 80 participants, with 40 in each condition. In
the preference condition, each table had a different object. In the
beginning, the animal appeared on the hill and introduced themselves.
Next, they turned to one of the tables and expressed either that they
liked (``Oh wow, I really like that one'') or disliked (``Oh bleh, I
really don't like that one'') the object before turning to the other
side and expressing the respective other attitude. Then, the animal
disappeared. After a short period of time, either the same or a
different animal appeared and requested an object while facing straight
ahead (see Fig. \ref{fig:design}, middle). If participants took into
account the information they gained about the speaker, they should
select the previously preferred object if the returning animal was the
same. If a different animal returned, they should choose randomly
between objects.

In the novelty condition, one table was initially empty while there was
an object on the other table (see Fig. \ref{fig:design}). The animal
turned to one of the sides and commented either on the presence (``Aha,
look at that'') or the absence of an object (``Hm, nothing there'').
Next, the animal disappeared. The same animal re-appeared and the
sequence above was repeated. When the animal disappeared for the second
time, a second object appeared on the empty table while the animal was
away. Like in the preference condition, we now manipulated if the same
animal or a different one returned. In case of the same animal
returning, listeners could infer the referent of the subsequent request
by considering that one object was new to the speaker and therefore more
likely to be of interest to them. However, no such inference was
licensed when a different animal returned because both objects were
novel.

@@

Participants selected the preferred object above chance when the same
animal returned (t(39) = 29.14, \emph{p} \textless{} .001, see Fig.
\ref{fig:plotexp12}) and did so more often compared to trials in which a
different animal returned (GLMM: \(\beta\) = 2.97, se = 0.7, \emph{p}
\textless{} .001). Thus, listeners inferred the referent of the
utterance by considering previous interactions with the speaker.
Interestingly, participants transferred preference to some extent from
one animal to the other and selected the preferred object above chance
when a different animal returned (t(39) = 2.7, \emph{p} = .01). In sum,
this study shows that adults make comparable inferences to children
(Saylor et al., 2009).

The novel object was selected above chance when the same animal returned
(t(39) = 6.77, \emph{p} \textless{} .001) but not when a different one
appeared (t(39) = 1.49, \emph{p} = .144, see Fig. \ref{fig:plotexp12}).
Furthermore, the two conditions differed in the expected direction
(\(\beta\) = 1.7, se = 0.39, \emph{p} \textless{} .001). Thus, like
children (Akhtar et al., 1996), adults used their prior information
about the speaker to resolve ambiguity in the utterance.

@@

\caption[Results from Experiment 3]{Results from Experiment 3. Dashed line indicates performance expected by chance. Plotting conventions are the same as in Fig. 2. All conditions in which CIs do not overlap with chance line are also statistically different from chance based on two-tailed Wilcoxon tests.}\label{fig:plotexp3}
\end{figure*}
\end{CodeChunk}

In Experiment 3, we combined the expectations studied in Experiment 1
and 2 to see how listeners integrate them.

@@

A total of 121 individuals participated in the experiment. The test
situation was the same as in the test condition in Experiment 1: One
table with object of type A and the other with an object of type A and
B. Again, the animal always turned to the table with two objects and
ambiguously requested an object. In Experiment 1, the listener had no
prior information about each object. In Experiment 3, however, we
manipulated common ground expectations in the same way as in Experiment
2. For example, the animal would turn to the table with one object and
express that they don't like object A, then turn to the other table and
express that they like object B. Next, after quickly disappearing, they
would reappear, turn to the table with two objects and make a request
(see Fig. \ref{fig:design}, right).

@@

To derive predictions, we used a probabilistic RSA model that simulates
a pragmatic listener reasoning about a cooperative speaker who is trying
to refer to an object (Frank \& Goodman, 2012). The speaker chooses how
to refer to the object by reasoning about a naive listener who does not
know the labels for the object (Frank \& Goodman, 2014). The conditional
probability that the listener chooses a referent given an utterance is
defined as follows: \[P_L(r_s|u)\propto P_S(u|r_s)P_S(r_s)\] Here,
\(P_S(u|r_s)\) is the likelihood that the speaker will use an utterance
u to refer to a specific referent r. It is defined in terms of a utility
function \(U_S(u;s)\) consisting of the surprisal of \(u\) for a naive
listener \(L_0\), who interprets \(u\) according its literal semantics:
\[P_S(u|r_s)\propto exp(\alpha U_S(u;s))\] The numerical strength of the
expression above depends on a scalar value, \(\alpha\), which can be
interpreted as an indicator of how rational the speaker is in choosing
utterances (i.e.~as \(\alpha\) increases, the speaker is more likely to
choose the most informative utterance).

@@

We used the results from Experiment 1 and 2 to specify \(\alpha\) as
well as \(P_S(r_s)\) in our model. We set \(\alpha\) so that a model
with uniform priors would predict the average proportion measured in
Experiment 1. The prior probability for each object was set to be the
proportion with which this object was chosen in Experiment
2\footnote{Proportions were measured when participants chose between two objects. However, in Experiment 3, three objects were involved. For each object we used the proportion measured in Experiment 2 as the prior probability.This approached spread out the absolute probability mass for each object but conserved the relative relation between objects.}.
Based on these parameter settings, we predicted the proportion with
which listeners will choose the more informative object in each of eight
unique conditions mentioned above (see also Fig. \ref{fig:plotexp3}). We
compared the fit of this pragmatic model to two alternative models using
Bayes Factors (BF). The first alternative model ignored the speaker
specific expectations (uniform prior model) while the second ignored the
informativeness inference (prior only model). All models included a
noise parameter, reflecting that participants may respond randomly
instead of in line with the intended manipulation on a given trial.
Noise parameters were estimated based on the data. They range between 0
and 1 and reflect the proportion of responses that are estimated to be
random instead of following the pattern predicted by the model.

@@

Results are discussed in the form of the proportion with which listeners
chose the more informative object (i.e., the object that would be the
more informative referent when only considering speaker general
information). For a comparison to chance within each unique condition
see \ref{fig:plotexp3}. Combinations of alignment and speaker identity
differed in how they influenced participants' responses (GLMM model
term: \texttt{alignment*speaker}; \(\beta\) = -2.92, se = 0.57, \emph{p}
\textless{} .001). Fig. \ref{fig:plotmodelcomp}A and B shows the mean
response in each unique condition compared to the pragmatics model.
Model predictions and data were highly correlated (r = 0.96, \emph{p}
\textless{} .001).Model fit was much better in the model taking into
account both types of expectations compared to the uniform prior (BF =
2e+79) or prior only model (BF = 1.8e+34). The inferred noise level in
the pragmatics model was 0.27 (95\% HDI: 0.21 - 0.34{]}.

Interestingly, as in Experiment 2, there was a transfer of preference in
the case of speaker change. Participants were at chance in the
preference - different speaker - incongruent condition (see Fig.
\ref{fig:plotexp3}). If preference would have been specific to a
particular individual, participants should have selected the less
frequent object above chance (as they did in the corresponding condition
with the novelty manipulation). Because it takes into account the
measurement from the earlier experiment, our model predicts these
results; future work might explicitly model generalization across
speakers.

@@

The strong manipulations were identical to Experiment 3 and the results
are therefore a direct replication (see Fig. \ref{fig:plotmodelcomp}C).
For novelty, in the medium condition, the animal turned to each table
only once before the test. In the weak condition, the animal only turned
to the table with an object before the test (instead of turning to and
commenting on both). In the medium condition for preference, the animal
only expressed liking and did so in a more subtle way (saying only:
``Oh, wow'' while pointing to the object). Participants were assigned to
one level of common ground expectation and completed two test trials in
each of the four conditions (alignment x speaker change).

@@

\caption[(A) Model predictions compared to data for Experiment 3 and (B) Experiment 4]{(A) Model predictions compared to data for Experiment 3 and (B) Experiment 4. (C) Data for strong prior manipulation in Experiment 3 and 4, providing a noise ceiling for the reliability of measurements. Error bars = 95\% HDIs.}\label{fig:plotmodelcomp}
\end{figure*}
\end{CodeChunk}

As noted above, the strong prior condition was a direct replication of
Experiment 3. Results from the two rounds of data collection were highly
correlated (r = 0.97, \emph{p} \textless{} .001, see Fig.
\ref{fig:plotmodelcomp}C). Across levels of prior manipulation, the data
from Experiment 4 were highly correlated to the corresponding model
predictions (r = 0.91, \emph{p} \textless{} .001, see Fig.
\ref{fig:plotmodelcomp}B). Again, the pragmatics model provided a much
better fit compared to the flat prior (BF = 4.4e+74) or prior only model
(BF = 1.8e+84). The inferred noise level in the pragmatics model was
0.28 (95\% HDI: 0.24 - 0.32{]}.

@@

Language use and learning requires balancing different types of
expectations about one's interlocutor - both expectations about how
speakers behave in general and specific expectations about how a
particular speaker might behave in a particular context. Here we used a
Bayesian pragmatics model to predict this integration process.
Experiment 1 and 2 replicated previous studies showing that adult
listeners expect speakers to produce utterances informatively and also
with respect to common ground. We then combined the procedures from the
first two experiments to study how listeners would integrate
expectations. We used the results from Experiment 1 and 2 to specify
model parameters that represented the two types of expectations,
generating predictions about new behavior. Experiments 3 and 4 showed
that both types of expectations influenced listeners inferences.
Overall, listener behavior was accurately described by our model,
suggesting that listeners trade-off flexibly between speaker specific
and general pragmatic expectations.

Notably, Experiment 3 also included situations in which the two
expectations were in conflict. For example, in some trials the speaker
expressed preference for object A, which was also the more frequent
(less informative) object. In these situations, a majority of
participants chose the preferred object as the referent (see also Fig.
\ref{fig:plotexp3} preference--same speaker--incongruent). A simple
explanation for this pattern might be that common ground manipulations
were simply ``stronger'', corroborated by the fact they produced higher
rates of expected choice than the informativeness expectation when the
two were presented in isolation (see Fig. \ref{fig:plotexp12}). In
Experiment 4, however, the medium manipulation for novelty yielded
numerically weaker results compared to the informativeness expectation
in Experiment 1, and yet participants still selected the novel object
above chance when the expectations were in conflict. Why is this?
Because common ground is represented in our model as the listener's
prior distribution, speakers can reason about it in choosing their
utterance. That is, in the mind of the listener, the speaker computes
the effect of each utterance on a naive listener with shared common
ground. Therefore, when prior interactions implicate one object as the
more likely referent, the speaker reasons that this object will be the
inferred referent of any semantically plausible utterance, even when the
same utterance would point to a different object in the absence of prior
information.

### 2019-01-29 stats/adults/ex3_2_variable_priors_comparison.Rmd
add kids so and pref models
first draft CogSci paper
add saves to cogsci folder

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

@@

bf <- data.frame(
  Comparison = c("Pragmatic vs. Prior only","Pragmatic vs. No prior", "No Prior vs. Prior only"),
  BayesFactor = c(
    exp(vp.llh.plot$logP[vp.llh.plot$Model== "Pragmatic"] - vp.llh.plot$logP[vp.llh.plot$Model== "Prior Only"]),
    exp(vp.llh.plot$logP[vp.llh.plot$Model== "Pragmatic"] - vp.llh.plot$logP[vp.llh.plot$Model== "No Prior"]),
    exp(vp.llh.plot$logP[vp.llh.plot$Model== "No Prior"] - vp.llh.plot$logP[vp.llh.plot$Model== "Prior Only"]))
)

bf%>%
  kable(digits = 2)
```

@@

@@

save.image("mcc_model_comparison.RData")
```

>>>>

```{r setup, include=FALSE}
load("../saves/mcc_model_comparison.RData")

knitr::opts_chunk$set(echo = TRUE)

@@

vp.bf <- data.frame(
  Comparison = c("Pragmatic vs. Prior only","Pragmatic vs. No prior", "No Prior vs. Prior only"),
  BayesFactor = c(
    exp(vp.llh.plot$logP[vp.llh.plot$Model== "Pragmatic"] - vp.llh.plot$logP[vp.llh.plot$Model== "Prior Only"]),
    exp(vp.llh.plot$logP[vp.llh.plot$Model== "Pragmatic"] - vp.llh.plot$logP[vp.llh.plot$Model== "No Prior"]),
    exp(vp.llh.plot$logP[vp.llh.plot$Model== "No Prior"] - vp.llh.plot$logP[vp.llh.plot$Model== "Prior Only"]))
)

vp.bf%>%
  kable(digits = 2)
```

@@

x <- p.llh.plot.variable %>%
  group_by(Model)%>%
  summarize(mean = mean(logP))

@@

save.image("../saves/mcc_model_comparison.RData")
```

### 2019-01-29 stats/kids/kids_novel.Rmd
add saves to cogsci folder

d_new <- read_csv(file="kids_novel.data.csv")%>%
  mutate(trial_type = ifelse(trial == "train", "train", "test"))

d_old <- read_csv(file="kids_novel.data.round1.csv")%>%
  mutate(trial_type = ifelse(trial == "train", "train", "test"))

>>>>

d_new <- read_csv(file="../data/kids_novel.data.csv")%>%
  mutate(trial_type = ifelse(trial == "train", "train", "test"))

d_old <- read_csv(file="../data/kids_novel.data.round1.csv")%>%
  mutate(trial_type = ifelse(trial == "train", "train", "test"))
